Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory

### 2024-10-25 paper-new/appendix.tex
paper revisions
latest changes

\section{Gradual Typing Syntax}

@@

\subsection{Free Composition of Error Domain Relations}
\begin{definition}\label{def:free-comp-ed-rel}
  Let $d : B_1 \rel B_2$ and $d' : B_2 \rel B_3$ be error domain relations. We
  define the composition $dd'$ inductively by the following rules:
  \begin{mathpar}
      \inferrule*[right = Comp]
      {b_1 \mathbin{d} b_2 \and b_2 \mathbin{d'} b_3}
      {b_1 \mathbin{d \relcomp d'} b_3}

      \inferrule*[right = DnClosed]
      {b_1' \le_{B_1} b_1 \and b_1 \mathbin{d \relcomp d'} b_3}
      {b_1' \mathbin{d \relcomp d'} b_3}

      \inferrule*[right = UpClosed]
      {b_1 \mathbin{d \relcomp d'} b_3 \and b_3 \le_{B_3} b_3'}
      {b_1 \mathbin{d \relcomp d'} b_3'}

      \inferrule*[right = PresErr]
      { }
      {\mho_{B_1} \mathbin{d \relcomp d'} b_3}

      \inferrule*[right = PresTheta]
      {\later_t( \tilde{b_1} \mathbin{d \relcomp d'} \tilde{b_3} ) }
      {\theta_{B_1}(\tilde{b_1}) \mathbin{d \relcomp d'} \theta_{B_3}(\tilde{b_3}) }
  \end{mathpar}

  This has the universal property that $\phi \ltsq{d \relcomp d'}{d''}
  \phi'$ if and only if for any $b \binrel{d} b'$ and $b' \binrel{d'}
  b''$ that $\phi(b) \binrel{d''} \phi'(b'')$.
\end{definition}

@@

>>>>

\section{Gradual Typing Syntax}\label{sec:appendix-gtlc-syntax}

@@

% \subsection{Free Composition of Error Domain Relations}
% \begin{definition}\label{def:free-comp-ed-rel}
%   Let $d : B_1 \rel B_2$ and $d' : B_2 \rel B_3$ be error domain relations. We
%   define the composition $dd'$ inductively by the following rules:
%   \begin{mathpar}
%       \inferrule*[right = Comp]
%       {b_1 \mathbin{d} b_2 \and b_2 \mathbin{d'} b_3}
%       {b_1 \mathbin{d \relcomp d'} b_3}

%       \inferrule*[right = DnClosed]
%       {b_1' \le_{B_1} b_1 \and b_1 \mathbin{d \relcomp d'} b_3}
%       {b_1' \mathbin{d \relcomp d'} b_3}

%       \inferrule*[right = UpClosed]
%       {b_1 \mathbin{d \relcomp d'} b_3 \and b_3 \le_{B_3} b_3'}
%       {b_1 \mathbin{d \relcomp d'} b_3'}

%       \inferrule*[right = PresErr]
%       { }
%       {\mho_{B_1} \mathbin{d \relcomp d'} b_3}

%       \inferrule*[right = PresTheta]
%       {\later_t( \tilde{b_1} \mathbin{d \relcomp d'} \tilde{b_3} ) }
%       {\theta_{B_1}(\tilde{b_1}) \mathbin{d \relcomp d'} \theta_{B_3}(\tilde{b_3}) }
%   \end{mathpar}

%   This has the universal property that $\phi \ltsq{d \relcomp d'}{d''}
%   \phi'$ if and only if for any $b \binrel{d} b'$ and $b' \binrel{d'}
%   b''$ that $\phi(b) \binrel{d''} \phi'(b'')$.
% \end{definition}

@@

% \subsection{Functorial Actions on Syntactic Perturbations}

% We provide details on the definition of the homomorphisms specifying the
% interpretation of perturbations corresponding for the functors $\li$, $U$,
% $\times$, and $\arr$.

% \begin{itemize}
%   \item To define the homomorphism $i_{\li A}$, by the universal property
%   of the coproduct of monoids it suffices to define two homomorphisms, one from
%   $\mathbb{N} \to \morbisimid{\li A}$ and one from $M_A \to \morbisimid{\li A}$.
%   The first homomorphism sends the generator $1 \in \mathbb{N}$ to the error
%   domain endomorpism $(\delta \circ \eta)^\dagger$, and the second sends $m_A \in
%   M_A$ to $\li(i_A(m_A))$ and we observe in both cases that the resulting
%   endomorphisms are bisimilar to the identity since the action of $\li$ on
%   morphisms preserves bisimilarity.

%   \item To define the homomorphism $i_{A_1 \times A_2}$, we
%   appeal to the universal property and give homomorphisms $M_{A_1} \to
%   \morbisimid{A_1 \times A_2}$ and $M_{A_2} \to \morbisimid{A_1 \times A_2}$. The
%   former applies $i_{A_1}$ in the first component and the identity in the second
%   component, while the latter does the reverse with $i_{A_2}$.

%   \item To define the homomorphism $i_{A \arr B}$ it suffices to define a
%   homomorphism $M_A^{op} \to \morbisimid{A \arr B}$ and $M_B \to \morbisimid{A
%   \arr B}$. The former takes an element $m_A$ and a predomain morphism $f : A \to
%   UB$ and returns the composition $f \circ i_A(m_A)$, while the latter is defined
%   similarly by post-composition with $i_B$.
